Drought regimens predict life history strategies in \emph{Heliophila}

### 2019-04-05 manuscript/manuscript.Rmd
update

abstract: \begin{center}\textbf{Summary}\end{center} \par Explaining variation in life history strategies is an enduring goal of evolutionary biology and ecology. Early theory predicted that for plants, annual and perennial life histories reflect adaptation to environments that experience alternative drought regimens. Nevertheless, empirical support for this hypothesis from comparative analyses remains lacking. \par Here, we test classic life history theory in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, controlling for phylogeny and integrating 34 years of satellite-based drought detection with 2,192 herbaria occurrence records. \par We find that the common ancestor of \textit{Heliophila} species was likely an annual, and that perenniality and annuality have repeatedly evolved, eight and five times, respectively. By comparing historical drought regimens, we show that annuals occur in environments where droughts are significantly more frequent than perennial species. We also provide evidence that annual plants adapt to predictable drought regimens by escaping drought prone seasons as seeds. \par These results yield compelling support for longstanding theoretical predictions by revealing the importance of drought frequency and predictability to explain plant life history. More broadly, this work highlights scalable approaches integrating herbaria records and remote sensing to address outstanding questions in evolutionary ecology.

@@

```
To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimens, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency.  First, we compared the frequency of drought across all occurrence records of between annual and perennial species by t-tests. Independent tests were performed to compare drought frequency between annuals and perennials during each season. We also tested for a relationship between drought frequency and life history, season, and their interaction by analysis of variance while including species as a random effect. To account for variation in the number of occurrence records per species, we next calculated the mean drought frequency during the winter, spring, summer and fall for each species.  Because shared evolutionary history of closely related species can lead to spurious associations between traits and environments  [@felsenstein1985phylogenies], we tested for a relationship between life history strategy and drought frequency while controlling for phylogeny using phylogenetic logistic regression [@ives2010phylogenetic]. This statistical approach is designed to control for the confounding effects of common ancestry's influence on demographic features such as geospatial relationships when addressing hypotheses about the role of natural selection on trait distributions.

@@

The topology of the estimated _Heliophila_ phylogeny was consistent with previous studies [@mummenhoff2005phylogeny;@mandakova2012whole]. Based on 10,000 stochastic character maps simulated under an equal rates model of character evolution in life history, an average of approximately eight changes from annual to perennial and five changes from perennial to annual are observed per stochastic character map (Figure \@ref(fig:phylogeny)a). These results suggest that the ancestral state of _Heliophila_ was annual and that both character states have arisen independently mutliple times.

>>>>

abstract: \begin{center}\textbf{Summary}\end{center} \par Explaining variation in life history strategies is an enduring goal of evolutionary biology and ecology. Early theory predicted that for plants, annual and perennial life histories reflect adaptation to environments that experience alternative drought regimens. Nevertheless, empirical support for this hypothesis from comparative analyses remains lacking. \par Here, we test classic life history theory in \textit{Heliophila} (Brassicaceae), a diverse genus of flowering plants native to Africa, controlling for phylogeny and integrating 34 years of satellite-based drought detection with 2,192 herbaria occurrence records. \par We find that the common ancestor of \textit{Heliophila} species was likely an annual, and that perenniality and annuality have repeatedly evolved, an estimated seven and five times, respectively. By comparing historical drought regimens, we show that annuals occur in environments where droughts are significantly more frequent than perennial species. We also provide evidence that annual plants adapt to predictable drought regimens by escaping drought prone seasons as seeds. \par These results yield compelling support for longstanding theoretical predictions by revealing the importance of drought frequency and predictability to explain plant life history. More broadly, this work highlights scalable approaches integrating herbaria records and remote sensing to address outstanding questions in evolutionary ecology.

@@

```
To evaluate the hypothesis that annual and perennial life history strategies reflect adaptations to alternative drought regimens, we tested the corresponding prediction that the observed distributions of annual and perennial _Heliophila_ species would be significantly associated with historic drought frequency. We  tested for a relationship between drought frequency and life history, season, and their interaction by analysis of variance while including species as a random effect using the lme4 package in R [@bates2014fitting] and compared annuals and perennials using by Tukey adjusted post-hoc contrasts. We next calculated the mean drought frequency during the winter, spring, summer and fall for each species.  Because shared evolutionary history of closely related species can lead to spurious associations between traits and environments  [@felsenstein1985phylogenies], we tested for a relationship between life history strategy and drought frequency while controlling for phylogeny using phylogenetic logistic regression [@ives2010phylogenetic]. This statistical approach is designed to control for the confounding effects of common ancestry's influence on demographic features such as geospatial relationships when addressing hypotheses about the role of natural selection on trait distributions.

@@

The topology of the estimated _Heliophila_ phylogeny was consistent with previous studies [@mummenhoff2005phylogeny;@mandakova2012whole]. Based on 10,000 stochastic character maps simulated under an equal rates model of character evolution in life history, an average of approximately seven changes from annual to perennial and five changes from perennial to annual are observed per stochastic character map (Figure \@ref(fig:phylogeny)a). These results suggest that the ancestral state of _Heliophila_ was annual and that both character states have arisen independently mutliple times.

### 2019-04-05 manuscript/manuscript.tex
update

We find that the common ancestor of \textit{Heliophila} species was
likely an annual, and that perenniality and annuality have repeatedly
evolved, eight and five times, respectively. By comparing historical
drought regimens, we show that annuals occur in environments where
droughts are significantly more frequent than perennial species. We also
provide evidence that annual plants adapt to predictable drought
regimens by escaping drought prone seasons as seeds.

@@

To evaluate the hypothesis that annual and perennial life history
strategies reflect adaptations to alternative drought regimens, we
tested the corresponding prediction that the observed distributions of
annual and perennial \emph{Heliophila} species would be significantly
associated with historic drought frequency. First, we compared the
frequency of drought across all occurrence records of between annual and
perennial species by t-tests. Independent tests were performed to
compare drought frequency between annuals and perennials during each
season. We also tested for a relationship between drought frequency and
life history, season, and their interaction by analysis of variance
while including species as a random effect. To account for variation in
the number of occurrence records per species, we next calculated the
mean drought frequency during the winter, spring, summer and fall for
each species. Because shared evolutionary history of closely related
species can lead to spurious associations between traits and
environments (Felsenstein, 1985), we tested for a relationship between
life history strategy and drought frequency while controlling for
phylogeny using phylogenetic logistic regression (Ives \& Garland,
2010). This statistical approach is designed to control for the
confounding effects of common ancestry's influence on demographic
features such as geospatial relationships when addressing hypotheses
about the role of natural selection on trait distributions.

@@

The topology of the estimated \emph{Heliophila} phylogeny was consistent
with previous studies (Mummenhoff \emph{et al.}, 2005; Mandáková
\emph{et al.}, 2012). Based on 10,000 stochastic character maps
simulated under an equal rates model of character evolution in life
history, an average of approximately eight changes from annual to
perennial and five changes from perennial to annual are observed per
stochastic character map (Figure \ref{fig:phylogeny}a). These results
suggest that the ancestral state of \emph{Heliophila} was annual and
that both character states have arisen independently mutliple times.

>>>>

We find that the common ancestor of \textit{Heliophila} species was
likely an annual, and that perenniality and annuality have repeatedly
evolved, an estimated seven and five times, respectively. By comparing
historical drought regimens, we show that annuals occur in environments
where droughts are significantly more frequent than perennial species.
We also provide evidence that annual plants adapt to predictable drought
regimens by escaping drought prone seasons as seeds.

@@

To evaluate the hypothesis that annual and perennial life history
strategies reflect adaptations to alternative drought regimens, we
tested the corresponding prediction that the observed distributions of
annual and perennial \emph{Heliophila} species would be significantly
associated with historic drought frequency. We tested for a relationship
between drought frequency and life history, season, and their
interaction by analysis of variance while including species as a random
effect using the lme4 package in R ({\textbf{???}}) and compared annuals
and perennials using by Tukey adjusted post-hoc contrasts. We next
calculated the mean drought frequency during the winter, spring, summer
and fall for each species. Because shared evolutionary history of
closely related species can lead to spurious associations between traits
and environments (Felsenstein, 1985), we tested for a relationship
between life history strategy and drought frequency while controlling
for phylogeny using phylogenetic logistic regression (Ives \& Garland,
2010). This statistical approach is designed to control for the
confounding effects of common ancestry's influence on demographic
features such as geospatial relationships when addressing hypotheses
about the role of natural selection on trait distributions.

@@

The topology of the estimated \emph{Heliophila} phylogeny was consistent
with previous studies (Mummenhoff \emph{et al.}, 2005; Mandáková
\emph{et al.}, 2012). Based on 10,000 stochastic character maps
simulated under an equal rates model of character evolution in life
history, an average of approximately seven changes from annual to
perennial and five changes from perennial to annual are observed per
stochastic character map (Figure \ref{fig:phylogeny}a). These results
suggest that the ancestral state of \emph{Heliophila} was annual and
that both character states have arisen independently mutliple times.
